---
title: 'Lab week 5: time series with feast & fable'
author: "Allison Horst"
date: "2/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

### 0. Attach packages

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(tsibble)
# library(tsibbledata)
library(feasts) # Possibly get dev version (gg_season issue): remotes::install_github("tidyverts/feasts")
library(forecast)
library(paletteer)
```

### 1. Get the data

We'll explore, then forecast, US energy consumption and production by renewables source. Get the data from `renewables_cons_prod.csv`:

```{r}
us_renew <- read_csv("renewables_cons_prod.csv") %>% 
  clean_names()
```

Explore the data frame:

```{r}
# View(us_renew)
# names(us_renew)
# unique(us_renew$description)
```

We'll focus on consumption data. 

### Clean up data

- Convert description to all lowercase
- Only keep observations for "consumption"
- Remove any "total" observations

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))
```

### Convert `yyyymm` column to tsibble with `lubridate`

```{r}
renew_date <- renew_clean %>% 
  mutate(month = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month = yearmonth(month)) %>% #coerce to `yearmonth` format
  mutate(value = as.numeric(value)) %>% 
  drop_na(month, value)
```

### Make a ggplot

Make, then save, the base line plot as `renew_gg`:

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month, y = value, group = description)) +
  geom_line(aes(color = description)) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 350))
```

Now try updating your color palette using options from paletteer. Use `View(palettes_d_names)` to see all of the discrete scale options. We'll want a palette with at least 7 colors (length >= 7). Find a name of a palette that you like, then update your graph by adding `scale_color_paletteer_d("package::palette")`. Like, if I want to use the `calecopal::figmtn` palette, I'd add:

`renew_gg + scale_color_paletteer_d("calecopal::figmtn")`

Try some out!

```{r}
renew_gg +
  scale_color_paletteer_d("calecopal::figmtn")
```

Have some fun trying out different color palettes.  

### Coerce to a tsibble:

```{r}
renew_ts <- as_tsibble(renew_date, key = description, index = month)
```

### Look at the data in a few different ways:
```{r}
renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
renew_ts %>% gg_season(value)
```

### Get just the wind energy consumption data:
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

# Explore: 
hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)

```

### Calculate summary data by time using `index_by()`

What if we want to calculate 

### Some hydro power consumption decomposition

First, let's check the decomposition (STL):
```{r}
# Find STL decomposition
dcmp <- hydro_ts %>%
  model(STL(value ~ season(window = Inf)))

# View the components
# components(dcmp)

# Visualize the decomposed components
components(dcmp) %>% autoplot() +
  theme_minimal()
```





<!---
The `fable` package "provides a collection of commonly used univariate and multivariate time series forecasting models including exponential smoothing via state space models and automatic ARIMA modelling". More information: https://github.com/tidyverts/fable

ARIMA forecasting: 
```{r, eval = FALSE}
wind_ts %>% model(
    ets = ETS(box_cox(value, 0.3)),
    arima = ARIMA(value),
  ) %>%
  forecast(h = "5 years") %>% 
  autoplot(filter(wind_ts, year(month) > 2010), level = NULL)

aus_retail %>%
  filter(
    State %in% c("New South Wales", "Victoria"),
    Industry == "Department stores"
  ) %>% 
  model(
    ets = ETS(box_cox(Turnover, 0.3)),
    arima = ARIMA(log(Turnover)),
    snaive = SNAIVE(Turnover)
  ) %>%
  forecast(h = "2 years") %>% 
  autoplot(filter(aus_retail, year(Month) > 2010), level = NULL)

# Testing:

df <- aus_retail %>%
  filter(
    State %in% c("New South Wales", "Victoria"),
    Industry == "Department stores"
  ) %>% 
  model(
    ets = ETS(box_cox(Turnover, 0.3)),
    arima = ARIMA(log(Turnover)),
    snaive = SNAIVE(Turnover)
  ) %>%
  forecast(h = "2 years")
```

---->


